---
title: "Join Easy Documentation (Rough Draft)"
output: html_document
---

This library depends on the `sqldf` library:
```{r message=FALSE}
library(sqldf)
```

Most functions in this library will follow the same pattern:
The first two arguments are data frames, and the third argument is a string specifying the primary key column's name.
If no primary key is specified, the row numbers will be used to create a primary key; this new column will be dropped at return.
The implementation and an example of the left join function is as follows:
```{r}
left_join <- function(left, right, key = NULL) {
  if (is.null(key)) {
    left$pk <- seq_len(nrow(left))
    right$pk <- seq_len(nrow(right))
    rtn <- sqldf("SELECT * FROM left l JOIN right r ON l.pk = r.pk")
    # TODO (DRY concerns): This should be extracted to its own, private function:
    rtn <- rtn[, names(rtn) != "pk"]
    return(rtn)
  }

  rtn <- fn$sqldf("SELECT * FROM left l JOIN right r ON l.$key = r.$key")
  # TODO (DRY concerns): This should also be extracted to its own, private function:
  return (rtn[, !duplicated(colnames(rtn))])
}

# Example usage without a primary key:
left1 <- data.frame(letter = c('a', 'b', 'c', 'd', 'e', 'f', 'g'))
right1 <- data.frame(number = c(1, 2, 3, 4, 5, 6, 7))
output1 <- left_join(left1, right1)
output1

# Example usage with a predefined primary key:
left2 <- data.frame(id = 100:106, letter = c('a', 'b', 'c', 'd', 'e', 'f', 'g'))
right2 <- data.frame(id = 100:106, number = c(1, 2, 3, 4, 5, 6, 7))
output2 <- left_join(left2, right2, key = "id")
output2
```

The left_join function (and similar) are comparable to R's merge function.
However, this library offers more complex operations such as that of the anti_join function:
```{r}
anti_join <- function(left, right, key = NULL) {
  if (is.null(key)) {
    left$pk <- seq_len(nrow(left))
    right$pk <- seq_len(nrow(right))
    rtn <- sqldf("SELECT * FROM left l WHERE NOT EXISTS (SELECT 1 FROM right r WHERE l.pk = r.pk)")
    return(rtn[, names(rtn) != "pk"])
  }

  rtn <- fn$sqldf("SELECT * FROM left l WHERE NOT EXISTS (SELECT 1 FROM right r WHERE l.$key = r.$key)")
  return (rtn[, !duplicated(colnames(rtn))])
}

# Example usage:
left3 <- data.frame(id = 100:106, letter = c('a', 'b', 'c', 'd', 'e', 'f', 'g'))
random_i <- sample(seq_len(nrow(left3)))
cutpoint <- round(nrow(left3) / 2)
right3 <- left3[random_i[seq_len(cutpoint)],]
output3 <- anti_join(left3, right3, key = "id")
output3
# right3 for comparison:
right3
```
